Resilience and Home-Space for WSTS \thanks{This work was partly done while the authors were supported by the Agence Nationale de la Recherche grant BraVAS (ANR-17-CE40-0028).}

### 2023-11-02 VMCAI-revised-version/final.tex
new simpler proof for undecidability

\begin{proof}
First, note that {\sc Resilience} is undecidable for a Minsky machine $M$ with at least $2$ counters when
$\Safe = \mathop{\downarrow} \Safe$,
since it is possible to have $\Safe = \{q_f(0,0)\}$. 
This stems from the fact that it is undecidable whether for a Minsky machine $M$ there exists a terminating run for every possible initial state.
\alain{et alors ? d'où sais-tu que le pb est indécidable pour safe=q,0 ? Bien sûr il l'est mais pas immédiatement puisque ce n'est pas le reachability}
Now, executions of Minsky machines can be simulated by reset-VASS~\cite{araki1976PN,dufourd1998reset} in such a way that any sequence 
% from  a state $p({\bf u})$
% to
reaching
 a state 
%
$ q({\bf 0})$ corresponds to one that reaches 
$q_f(0,0)$
 in the Minsky machine and vice versa.
 The construction rely on using pairs of counters to represent counters of the Minsky machine,
where instead of being tested for zero before a transition, one of the the counters in the pair is being reset; if the counters were zero before the transition then this was a faithful simulation, 
otherwise, there is a gap between the two counters which cannot be undone otherwise than by letting the machine backtrack.
Using the construction, which can be seen in more details in Appendix~\ref{minsk},
 we obtain that 
 {\sc Resilience} is undecidable for reset-VASS when $\Safe = \downarrow \Safe$, and hence for  $\post$-effective, $\pred$-basis-effective, 
$\omega^2$-WSTS with strong compatibility as well.\qed \end{proof}

@@

\begin{proof}
Let $V$ be a reset-VASS and $q_0({\bf 0})$ an initial state of $V$,
we build $V'$ from $V$ by additionnally allowing from any control-state to get back to 
$q_0({\bf 0})$
by a transition reseting every counter and
changing the control-state to $q_0$.
Then a state $q_f({\bf 0})$ is reachable from $q_0({\bf 0})$ in $N$
iff $q_f({\bf 0})$ is reachable from every element of $\post^*(q_0({\bf 0}))$ in $N'$,
which corresponds to {\sc State-resilience}  for $\Safe =  \{ q_f({\bf 0}) \}$
and $s_0 = q_0({\bf 0})$.
\end{proof}

@@

\section{Reset-VASS and Minsky-machines}\label{minsk}

We now delve into more details into the undecidability results 
upon which rely
Proposition~\ref{indec WSTS with dcs}
and
Proposition~\ref{srp down up},
namely, that
{\sc Resilience} and {\sc State-resilience} when $\Safe = \mathop{\downarrow}~\Safe$
are undecidable for Minsky machines
and 
for reset-VASS.

The problems are undecidable for a Minsky machine with at least $2$ counters when
$\Safe$ is downward-closed, since it is possible to consider $\Safe = \{q_f(0,0)\}$. We then use Reset-VASS to simulate the Minsky machines~\cite{araki1976PN,dufourd1998reset}, albeit in a weak way. 
% The simulation however is 
Since by a particular choice of $\Safe$ and $\Bad$ we are only interested in the reachability of $q_f(0,0)$ in the Minsky machine, 
this weak simulation will be enough to obtain undecidability of both problems.

@@

Now, executions of Minsky machines can be simulated by reset-VASS~\cite{araki1976PN,dufourd1998reset}. 
% \alain{\cite{araki1976PN} simule-t-il les Minsky machines et comment ?}
%Reset-VASS extend the basic VASS model with special “reset
%transitions” that set to $0$ some coordinates in the vector. 
%			Let us recall their definition here.
%      	it is well known that reset-VASS are WSTS~\cite{dufourd1998reset}.
The simulation operates by building a reset-VASS 
which mimick the behaviors of the Minsky machines' counters by using two mirror counters 
for each of them,
and where instead of being tested for zero before a transition, one of the the counters in the pair is being reset; if the counters were zero before the transition then this was a faithful simulation, 
otherwise, there is a gap between the two counters which cannot be undone otherwise than by letting the machine backtrack.
The construction is such that
there exists
 a sequence from 
an initial state %$(s,x_1 \ldots x_n)$
	with equal mirror counters
to
a state $ q({\bf 0})$ if and only if
there exists
a sequence
 in the Minsky machine
 from a corresponding initial state
 to
 the state $q_f(0,0,0)$
  (from Theorem~$5$ in \cite{araki1976PN}).
  [... from initial states with unequal mirror counters ... -> go back to a state with equal mirror counters -> then ...]
Hence {\sc Resilience} (resp. {\sc State resilience}) holds in the Minsky machine $M'$ iff 
{\sc Resilience} (resp. {\sc State resilience}) holds in the reset-VASS simulating $M'$.
Since reset-VASS can simulate executions of a Minsky machine, {\sc Resilience} 
and  {\sc State-resilience} are undecidable for reset-VASS and hence for  $\pred$-basis-effective, 
$\omega^2$-WSTS with strong compatibility as well.\qed
\end{proof}

>>>>

\begin{proof}
Let $V$ be a reset-VASS and $q_0({\bf 0})$ an initial state of $V$,
we build $V'$ from $V$ by additionnally allowing from any control-state to get back to 
$q_0({\bf 0})$
by a transition reseting every counter and
changing the control-state to $q_0$.
Then a state $q_f({\bf 0})$ is reachable from $q_0({\bf 0})$ in $N$
iff $q_f({\bf 0})$ is reachable from every element in $V'$,
which corresponds to {\sc Resilience}  for $\Safe =  \{ q_f({\bf 0}) \}$.
\end{proof}

	\iffalse
\begin{proof}
First, note that {\sc Resilience} is undecidable for a Minsky machine $M$ with at least $2$ counters when
$\Safe = \mathop{\downarrow} \Safe$,
since it is possible to have $\Safe = \{q_f(0,0)\}$. 
This stems from the fact that it is undecidable whether for a Minsky machine $M$ there exists a terminating run for every possible initial state.
\alain{et alors ? d'où sais-tu que le pb est indécidable pour safe=q,0 ? Bien sûr il l'est mais pas immédiatement puisque ce n'est pas le reachability}
Now, executions of Minsky machines can be simulated by reset-VASS~\cite{araki1976PN,dufourd1998reset} in such a way that any sequence 
% from  a state $p({\bf u})$
% to
reaching
 a state 
%
$ q({\bf 0})$ corresponds to one that reaches 
$q_f(0,0)$
 in the Minsky machine and vice versa.
 The construction rely on using pairs of counters to represent counters of the Minsky machine,
where instead of being tested for zero before a transition, one of the the counters in the pair is being reset; if the counters were zero before the transition then this was a faithful simulation, 
otherwise, there is a gap between the two counters which cannot be undone otherwise than by letting the machine backtrack.
Using the construction, which can be seen in more details in Appendix~\ref{minsk},
 we obtain that 
 {\sc Resilience} is undecidable for reset-VASS when $\Safe = \downarrow \Safe$, and hence for  $\post$-effective, $\pred$-basis-effective, 
$\omega^2$-WSTS with strong compatibility as well.\qed \end{proof}
	\fi

@@

\begin{proof}
Let $V$ be a reset-VASS and $q_0({\bf 0})$ an initial state of $V$,
we build $V'$ from $V$ by additionnally allowing from any control-state to get back to 
$q_0({\bf 0})$
by a transition reseting every counter and
changing the control-state to $q_0$.
Then a state $q_f({\bf 0})$ is reachable from $q_0({\bf 0})$ in $N$
iff $q_f({\bf 0})$ is reachable from every element of $\post^*(q_0({\bf 0}))$ in $V'$,
which corresponds to {\sc State-resilience}  for $\Safe =  \{ q_f({\bf 0}) \}$
and $s_0 = q_0({\bf 0})$.
\end{proof}

@@

	\iffalse
\section{Reset-VASS and Minsky-machines}\label{minsk}

We now delve into more details into the undecidability results 
upon which rely
Proposition~\ref{indec WSTS with dcs},
% and
% Proposition~\ref{srp down up},
namely, that
{\sc Resilience} %and {\sc State-resilience} 
when $\Safe = \mathop{\downarrow}~\Safe$
%are
is
%
 undecidable for Minsky machines
and 
for reset-VASS.

The 
% problems are
problem is undecidable for a Minsky machine with at least $2$ counters when
$\Safe$ is downward-closed, since it is possible to consider $\Safe = \{q_f(0,0)\}$. We then use Reset-VASS to simulate the Minsky machines~\cite{araki1976PN,dufourd1998reset}, albeit in a weak way. 
% The simulation however is 
Since by a particular choice of $\Safe$ and $\Bad$ we are only interested in the reachability of $q_f(0,0)$ in the Minsky machine, 
this weak simulation will be enough to obtain undecidability of both problems.

@@

Now, executions of Minsky machines can be simulated by reset-VASS~\cite{araki1976PN,dufourd1998reset}. 
% \alain{\cite{araki1976PN} simule-t-il les Minsky machines et comment ?}
%Reset-VASS extend the basic VASS model with special “reset
%transitions” that set to $0$ some coordinates in the vector. 
%			Let us recall their definition here.
%      	it is well known that reset-VASS are WSTS~\cite{dufourd1998reset}.
The simulation operates by building a reset-VASS 
which mimick the behaviors of the Minsky machines' counters by using two mirror counters 
for each of them,
and where instead of being tested for zero before a transition, one of the the counters in the pair is being reset; if the counters were zero before the transition then this was a faithful simulation, 
otherwise, there is a gap between the two counters which cannot be undone otherwise than by letting the machine backtrack.
The construction is such that
there exists
 a sequence from 
an initial state %$(s,x_1 \ldots x_n)$
	with equal mirror counters
to
a state $ q({\bf 0})$ if and only if
there exists
a sequence
 in the Minsky machine
 from a corresponding initial state
 to
 the state $q_f(0,0,0)$
  (from Theorem~$5$ in \cite{araki1976PN}).
  [... from initial states with unequal mirror counters ... -> go back to a state with equal mirror counters -> then ...]
  [ initial states with unequal mirror counters should be allowed to have some kind of backtracking to them]
Hence {\sc Resilience} (resp. {\sc State resilience}) holds in the Minsky machine $M'$ iff 
{\sc Resilience} (resp. {\sc State resilience}) holds in the reset-VASS simulating $M'$.
Since reset-VASS can simulate executions of a Minsky machine, {\sc Resilience} 
and  {\sc State-resilience} are undecidable for reset-VASS and hence for  $\pred$-basis-effective, 
$\omega^2$-WSTS with strong compatibility as well.\qed
\end{proof}
